Multiple testing with logical dependencies

### 2013-08-15 paper/draft.tex
fixed theorem after yet another false proof

\begin{theorem}
  Consider a system of logically dependent hypotheses $\mathscr{H}$
  with dependence relation $\sim$ and associated matroid
  $M(\mathscr{H})$. Assume that $M(\mathscr{H})$ neither contains
  loops nor parallels. Consider a hierarchical testing procedure that
  tests $H_{i}$ at level $\alpha$ according to a prefixed
  order. Without loss of generality assume that $H_i$ are indexed
  according to the fixed rejection sequence $(1,...,n)$. Given 
  that hypotheses with indeces $1$ through $r$ have been
  rejected and if for $r+1$ to $r+s$ holds
  \begin{equation}
    \label{eq:cond.hierarchical}
    \rho \{1,...,r\} + 1 = \rho \{1,...,r+1\} = \rho \{1,...,r+s\},
  \end{equation}
  hypotheses $H_{r+1}$ through $H_{r+s}$ can be tested simultaneously
  at full level $\alpha$.
\end{theorem}

\begin{proof}
  We show that given $p_{r+k} \leq \alpha$ for some $1 \leq k \leq s$
  all exhaustive sets $E \in \mathscr{L}_{r+k} = \{E \in \mathscr{L}:
  r+k \in E\}$ are rejected by the stepwise procedure. It remains to
  be shown that the stepwise procedure also rejects $H_E$ for all
  exhaustive sets with $\min E < r+k$. Note that $\{E \in \mathscr{L}:
  \min E < r+k\} = \overline{\underline{E}}_1^r \cup
  \overline{\underline{E}}^{r+k-1}_{r+1}$ where
  $\overline{\underline{E}}^b_a := \{\sigma I: \text{for all } I
  \subseteq S, a \leq \min I \leq b\}$. For $ E \in
  \overline{\underline{E}}_1^r$ follows since $p_i \leq \alpha$ for $i
  \leq r$.

  We show by contradiction that we have $i \in \sigma \{r+l,r+k\}$ for
  some $i \leq r$ and all $1 \leq l < k$.  Assume that there is an $l$
  such that there is no $i \leq r$ with $i \in \sigma
  \{r+l,r+k\}$. Consequently for all $i \leq r$, we have $i \nsim
  \{r+l,r+k\}$. Since $M(\mathscr{H})$ has no loops or parallels also
  $(r+l) \nsim (r+k)$. Consequently, for any maximally independent set
  $I \subset \{1,...,r\}$, $I \cup \{r+l, r+k\}$ is independent and
  has cardinality $\rho\{1,...,r\}+2$ which contradicts
  (\ref{eq:cond.hierarchical}).

  Above argument can be repeated for any set $I$ with $r+1 \leq \min I
  < k$. Therefore, we have $i \in E$ for all $E \in
  \overline{\underline{E}}^{r+k-1}_{r+1}$. Rejection of these sets
  follows from rejection of $H_i$, $i \leq r$, which completes the
  proof. 
\end{proof}

>>>>

\begin{theorem}
  Consider a system of logically dependent hypotheses $\mathscr{H}$
  with dependence relation $\sim$ and associated matroid
  $M(\mathscr{H})$. Assume that $M(\mathscr{H})$ neither contains
  loops nor parallels. Consider a hierarchical testing procedure that
  tests $H_{i}$ at level $\alpha$ according to a prefixed
  order. Without loss of generality assume that $H_i$ are indexed
  according to the fixed rejection sequence $(1,...,n)$. Given 
  that hypotheses with indeces $1$ through $r$ have been
  rejected and if for $r+1$ to $r+s$ holds
  \begin{equation}
    \label{eq:cond.hierarchical}
    \forall 1 \leq l < k \leq s: \; \exists i \leq r: i \sim \{l,k\},
  \end{equation}
  hypotheses $H_{r+1}$ through $H_{r+s}$ can be tested simultaneously
  at full level $\alpha$.
\end{theorem}

\begin{proof}
  We show that given $p_{r+k} \leq \alpha$ for some $1 \leq k \leq s$
  all exhaustive sets $E \in \mathscr{L}_{r+k} = \{E \in \mathscr{L}:
  r+k \in E\}$ are rejected by the stepwise procedure. Since $p_{r+k}
  \leq \alpha$ implies rejection of all $E$ such that $\min E \geq
  r+k$, it remains to be shown that the stepwise procedure rejects
  $H_E$ for all exhaustive sets with $\min E < r+k$. Note that $\{E
  \in \mathscr{L}: \min E < r+k\} = \overline{\underline{E}}_1^r \cup
  \overline{\underline{E}}^{r+k-1}_{r+1}$ where
  $\overline{\underline{E}}^b_a := \{\sigma I: \text{for all } I
  \subseteq S, a \leq \min I \leq b\}$. For $ E \in
  \overline{\underline{E}}_1^r$ follows since $p_i \leq \alpha$ for $i
  \leq r$.

  Consider any set $I$ with $r < \min I < k$ then there is an $i \sim
  \{\min I,k\}$ with $i \leq r$. Consequently, $i \in \sigma \{\min
  I,k\}$ and also $\sigma I$. Therefore, we have $i \in E$ for all $E
  \in \overline{\underline{E}}^{r+k-1}_{r+1}$. Rejection of these sets
  follows from rejection of $H_i$, $i \leq r$, which completes the
  proof.
\end{proof}
